Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-10-24 crowdsource/article.tex
Put back Pierre's explanation
Add the description of x^t_i

Under the simplifying assumption that neurons are on-off units, characterised
by short periods of intense activity, or peaks, and longer periods of
inactivity, the first part of our algorithm consists in cleaning the raw
fluorescence data. This pre-processing applies successively four
filtering functions to the original signal:
\begin{itemize}
  \item The first function smooths the fluorescence signal, using one of the following
low-pass filters for filtering out high frequency noise:
\begin{align} f_1(x^t_i) &= x^{t-1}_i + x^t_i + x^{t+1}_i, \label{eq:symetric-median} \\
  f_2(x^t_i) &= 0.4 x^{t-3}_i + 0.6 x^{t-2}_i + 0.8 x^{t-1}_i + x_i^t.\label{eq:weighted-asymetric-median}\\
  f_3(x^t_i) &= x^{t-1}_i + x^{t}_i + x^{t+1}_i + x^{t+2}_i, \label{eq:asymetric-median-forward} \\
f_4(x^t_i) &=  x_i^t + x^{t+1}_i  + x^{t+2}_i + x^{t+3}_i. \label{eq:asymetric-median}
\end{align}
\item To have a signal that only has high magnitude around instances where the
spikes occur, the second filter function transforms the time-series into
its backward difference 
\begin{align} g(x^{t}_{i}) &= x^{t}_i - x^{t-1}_i.
\label{eq:high-pass-filter}
\end{align}
\item To filter out small variations in the signal obtained after applying the
function $g$, as well as to eliminate negative values, we then use the following
hard-threshold filter
\begin{align}\label{eqn:hfilter}
h(x^{t}_i) &= (x^{t}_i)^c \mathbb{1}(x^{t}_i \geq \tau) \text{ with } \tau > 0,
\end{align}
where $\tau$ is the threshold parameter, $c$ is equal to 0.9, and $\mathbb{1}$ is the indicator
function.
\item The last filtering functions is defined as:
\begin{align}
 w^*(x^{t}_i) &= {(x^{t}_i + 1 )^{\left (1 + \frac{1}{\sum_{j} x^{t}_j}\right )}}^{k(\sum_{j} x^{t}_j)}
\end{align}
where the function $k$ is a piecewise linear function optimised
separately for each filter $f_1$, $f_2$, $f_3$ and $f_4$ (see the
implementation for full details). The effect of this filtering is to
magnify the importance of spikes that occur in cases of low global
activity (measured by $\sum_j x^t_j$).
\end{itemize}
The pre-processed times-series are then obtained by the application of
the following function: $w^*\circ h \circ g \circ f_i$ (with
$i=1$, 2, 3, or 4).

@@

\paragraph{Edge orientation.}

Partial correlation is a symmetric measure that can not be used to predict edge
orientation. A post-processing heuristic was proposed to derive an asymmetric
association score from the partial correlation coefficient. This heuristic is
based on the observation that if there is directed edges from $i$ to $j$ then
the activation of $i$ should be followed by the activation of $j$ after a
slight delay. For every pairs $(i,j)$, we computed the following term:
\begin{align}
s_{i,j} = \sum_{t=1}^{T - 1} \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2\right])
\end{align}
that could be interpreted as an image of the number of times that neuron $i$
activates neuron $j$. $\phi_1$ and $\phi_2$ are parameters whose values have
been chosen equal to $0.2$ and $0.5$, respectively. The
final association score for an edge from $i$ to $j$ was
then the following:
\begin{align}
q_{i,j} = weight * p_{i,j} + (1-weight) * (s_{i,j}-s_{j,i}),
\label{eqn:qij}
\end{align}
with $p_{i,j}$ the average partial correlation coefficient and the parameter
$weight$ chosen close to 1 ($weight=0.997$). The second term is expected to be
greater (resp. lower) than 0 when $i$ more often activates $j$ than the
opposite and therefore in this case it makes $q_{i,j}$ greater (resp. lower)
than $q_{j,i}$.

\newpage
\bibliography{references}

>>>>

Under the simplifying assumption that neurons are on-off units, characterised
by short periods of intense activity, or peaks, and longer periods of
inactivity, the first part of our algorithm consists in cleaning the raw
fluorescence data. This pre-processing applies successively four
filtering functions to the original signal $x^t_i$ of each neuron $i$ where $t$ is the time step:
\begin{itemize}
  \item The first function smooths the fluorescence signal, using one of the following
low-pass filters for filtering out high frequency noise:
\begin{align} f_1(x^t_i) &= x^{t-1}_i + x^t_i + x^{t+1}_i, \label{eq:symetric-median} \\
  f_2(x^t_i) &= 0.4 x^{t-3}_i + 0.6 x^{t-2}_i + 0.8 x^{t-1}_i + x_i^t.\label{eq:weighted-asymetric-median}\\
  f_3(x^t_i) &= x^{t-1}_i + x^{t}_i + x^{t+1}_i + x^{t+2}_i, \label{eq:asymetric-median-forward} \\
f_4(x^t_i) &=  x_i^t + x^{t+1}_i  + x^{t+2}_i + x^{t+3}_i. \label{eq:asymetric-median}
\end{align}
\item To have a signal that only has high magnitude around instances where the
spikes occur, the second filter function transforms the time-series into
its backward difference 
\begin{align} g(x^{t}_{i}) &= x^{t}_i - x^{t-1}_i.
\label{eq:high-pass-filter}
\end{align}
\item To filter out small variations in the signal obtained after applying the
function $g$, as well as to eliminate negative values, we then use the following
hard-threshold filter
\begin{align}\label{eqn:hfilter}
h(x^{t}_i) &= (x^{t}_i)^c \mathbb{1}(x^{t}_i \geq \tau) \text{ with } \tau > 0,
\end{align}
where $\tau$ is the threshold parameter, $c$ is equal to 0.9, and $\mathbb{1}$ is the indicator
function.
\item The last filtering functions is defined as:
\begin{align}
 w^*(x^{t}_i) &= {(x^{t}_i + 1 )^{\left (1 + \frac{1}{\sum_{j} x^{t}_j}\right )}}^{k(\sum_{j} x^{t}_j)}
\end{align}
where the function $k$ is a piecewise linear function optimised
separately for each filter $f_1$, $f_2$, $f_3$ and $f_4$ (see the
implementation for full details). The effect of this filtering is to
magnify the importance of spikes that occur in cases of low global
activity (measured by $\sum_j x^t_j$).
\end{itemize}
The pre-processed times-series are then obtained by the application of
the following function: $w^*\circ h \circ g \circ f_i$ (with
$i=1$, 2, 3, or 4).

@@

\paragraph{Edge orientation.}
Partial correlation is a symmetric measure that can not be used to predict edge
orientation. A post-processing heuristic was proposed to derive an asymmetric
association score from the partial correlation coefficient. This heuristic is
based on the observation that if there is directed edges from $i$ to $j$ then
the activation of $i$ should be followed by the activation of $j$ after a
slight delay. For every pairs $(i,j)$, we computed the following term:
\begin{align}
s_{i,j} = \sum_{t=1}^{T - 1} \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2\right])
\end{align}
that could be interpreted as an image of the number of times that neuron $i$
activates neuron $j$. $\phi_1$ and $\phi_2$ are parameters whose values have
been chosen equal to $0.2$ and $0.5$, respectively. The
final association score for an edge from $i$ to $j$ was
then the following:
\begin{align}
q_{i,j} = weight * p_{i,j} + (1-weight) * (s_{i,j}-s_{j,i}),
\label{eqn:qij}
\end{align}
with $p_{i,j}$ the average partial correlation coefficient and the parameter
$weight$ chosen close to 1 ($weight=0.997$). The second term is expected to be
greater (resp. lower) than 0 when $i$ more often activates $j$ than the
opposite and therefore in this case it makes $q_{i,j}$ greater (resp. lower)
than $q_{j,i}$.
\newpage
\bibliography{references}
